Hypertension and microbiome

### 2019-12-20 articletwo.Rmd
Init

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
```

@@

### Officer functions

<details>
  <summary>Open/Close</summary>

```{r Officer importer, code=readLines("articletwo-officer.R")}
source("articletwo-officer.R")
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

### Plot functions

<details>
  <summary>Open/Close</summary>

```{r Plot functions, code=readLines("articletwo-ggplot.R")}
source("articletwo-ggplot.R")
```

@@

@@

```{r Characteristics}
tbl1 <- tableone(meta(pseq.species))
tbl1
```

@@

Comparing how adding additional covariate to basic model for age, sex,
and hypertension changes number of significant associations found for
hypertension. Full model has

`r sum(results(dds$HYPERTENSION, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`

associations for hypertension.

@@

```{r deseq heatmap}
g.deseq <- deseqhaetmap(dset.deseq.signf)
ggsave(file = "cache/deseq.png", plot = g.deseq, height = 3, width = 12, unit = "in", dpi = 300)
```

@@

```{r salt phyloseq object}
rownames.salt <- pseq.genus %>% meta %>% subset(!is.na(NA.)) %>% rownames
pseq.genus.salt <- pseq.genus %>%
    prune_samples(rownames.salt, .) %>%
    prune_taxa(taxa(pseq.genus.core), .)
dset.genus.salt <- prune_lactobacillus(pseq.genus.salt, transform = "clr")
```

General linear models for associations between dU-NA and hypertension indices

```{r salt hypertension}
ret.salt <- calculateglm(dset.genus.salt, covariates = c("NA.", var.CL), filterstr = "NA.")
ret.salt %>% select(response, term, estimate, p.value) %>% kable
```

```{r salt characteristics}
dset.genus.salt %>%
    mytableone() %>%
    kable
```

```{r lactobacillus prevalence}
prune_lactobacillus(pseq.genus.salt, drop = FALSE) %>%
    summarize(lacto_detected = sum(lacto.prevalent),
              n = n(),
              ratio = lacto_detected/n) %>%
    kable
```

```{r lacto scatter plot}
g.salt <- saltplot(dset.genus.salt)
ggsave(file = "cache/gsalt.png", plot=g.salt, height=4, width=4, dpi = 300)
```

Deseq for lactobacillus

@@

```{r lacto deseq diagnostic}
png("cache/salt-ma.png", width = 1200, height = 800, pointsize = 20)
par(mfrow=c(1,2))
plotMA(dds.salt)
plotDispEsts(dds.salt)
dev.off()
```

<img src = "cache/salt-ma.png" />

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
!library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
```

@@

<details><summary>Functions</summary>

```{r Functions}
sourcefiles <- c("articletwo-officer.R",
                 "articletwo-rrbiome.R",
                 "articletwo-ggplot.R")
```

```{r import files, echo = FALSE}
for (f in sourcefiles) {
    source(f)
}
```

```{r embbed files, echo = FALSE}
xfun::embed_files(c("rrnmr.Rmd", sourcefiles))
```

@@

Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
genus `r pseq.genus %>% sample_sums %>% mean`.

@@

```{r Characteristics}
tbl1 <- tableone(meta(pseq.species))
```

@@

Comparing how adding additional covariate to basic model for age, sex,
and hypertension changes number of significant associations found for
hypertension.

@@

```{r deseq heatmap}
g.deseq <- deseqhaetmap(dset.deseq.signf)
ggsave(file = "cache/deseq.png", plot = g.deseq, height = 8, width = 4, unit = "in", dpi = 300)
```

@@

@@

```{r scatter plot from deseq}
df.mu <- assays(dds.salt)[["mu"]] %>%
    as.data.frame %>%
    tibble::rownames_to_column("rowname") %>%
    gather(sampleid, value, -rowname) %>%
    spread(rowname, value)

df.countmu <- full_join(meta(pseq.genus.salt) %>% tibble::rownames_to_column("sampleid"),
                        df.mu %>% select(sampleid, g_Lactobacillus),
                        by = "sampleid")

g.salt <- saltplot(df.countmu)
ggsave(file = "cache/gsalt.png", plot=g.salt, height=4, width=4, dpi = 300)
```
